|

### 2022-11-13 manuscript.Rmd
address reviewer comments

Another important aspect in curriculum design includes the timing of when to introduce accessibility concepts in data science courses. For example, some instructors may prefer to put teaching accessibility on the back burner so that they can save more time for their students to learn data science fundamentals first and adopt accessibility principles in an upper-level course later. While the decision is subject to each institution's and individual instructor's discussion, we believe the sooner, the better. This is because: (1) some students may take an introductory level only, which means they will lose a chance to get exposed to this demanding knowledge unless they elect the upper-level course covering accessibility; (2) this kind of separation may unintentionally contribute to misconceptions, such as accessibility is difficult and/or accessibility is optional, not essential. When properly incorporated, accessibility can be taught along with other data science fundamentals (e.g., wrangling, merging, tidying, visualizing) while not necessarily putting more cognitive burden against students' current learning. Rather, students can take a chance to become creative design thinkers by making their data science work more meaningful and accessible through "empathy learning" [@putnamTeachingAccessibilityLearning2015]. Thus, we suggest teaching accessibility as part of an introductory data science course for students to practice coupling their data science knowledge with accessibility principles as early as possible.

@@

We believe that a natural connection between data science and accessibility can be achieved in the data science classroom while presenting data visualization and then introducing the other aforementioned forms of data representation.
While delivering the content, instructors should try to provide as many opportunities as possible for students to experience using these tools themselves. 
For instance, each student should use a screen reader at least once regardless of their visual ability status and hear the alternate text that they have written.

Ample opportunities in learning accessibility should also include assessments. 
For instance, instructors incorporating accessibility should modify assessment instructions and rubrics to include accessibility. 
This will prevent students from just hearing about accessibility in the lecture and then forgetting about it. 
Ideally, the assessments should not treat accessibility as a learning objective at a single point in the academic term. 
For instance, if accessibility is covered in Week 3 of the term, students can (and should) still be expected to write alternate texts for visualizations in Week 10.

@@

Although we have mainly focused on tools and different representations of data to have students develop accessible content, a more crucial accessibility feature is reproducible workflows. 
Scholars often focus on the reproducibility and replication of scientific findings. 
We believe reproducibility can often serve as an accessibility feature in a data science context. 
A reproducible open-source workflow allows users who are blind or have a visual impairment to examine data in their preferred form of data representation.
Thus, in addition to teaching different representations of data, instructors may also consider teaching reproducibility.

@@

We believe that an accessible future in data science requires educators to teach accessibility as part of the curriculum deliberately. 
We also acknowledge that what we have presented here is only a small portion of the intersection of visual accessibility and data science. 
The second co-author was supported through a Teach Access grant to develop curricular materials on teaching accessibility as part of the data science curriculum. 
Educators interested in teaching accessibility can follow the organization's work at https://teachaccess.org/. 
The reproducible R source code of this manuscript can be found on a GitHub repository at \url{\repo{}}.

>>>>

Another important aspect of curriculum design includes the timing of when to introduce accessibility concepts in data science courses. For example, some instructors may prefer to put teaching accessibility on the back burner so that they can save more time for their students to learn data science fundamentals first and adopt accessibility principles in an upper-level course later. While the decision is subject to each institution's and individual instructor's discussion, we believe the sooner, the better. This is because: (1) some students may take an introductory level only, which means they will lose a chance to get exposed to this demanding knowledge unless they elect the upper-level course covering accessibility; (2) this kind of separation may unintentionally contribute to misconceptions, such as accessibility is difficult and/or accessibility is optional, not essential. When properly incorporated, accessibility can be taught along with other data science fundamentals (e.g., wrangling, merging, tidying, visualizing) while not necessarily putting more cognitive burden on students' current learning. Rather, students can take a chance to become creative design thinkers by making their data science work more meaningful and accessible through "empathy learning" [@putnamTeachingAccessibilityLearning2015]. Thus, we suggest teaching accessibility as part of an introductory data science course for students to practice coupling their data science knowledge with accessibility principles as early as possible.

@@

We believe that a natural connection between data science and accessibility can be achieved in the data science classroom while presenting data visualization and then introducing the other aforementioned forms of data representation.
While delivering the content, instructors should try to provide as many opportunities as possible for students to experience using these tools and accessibility principles themselves. 
For instance, each student could use screen reader at least once regardless of their visual acuity and hear the alternate text that they have written.

Ample opportunities in learning accessibility should also include assessments. 
For instance, instructors should modify assessment instructions and rubrics to include accessibility. 
This will prevent students from just hearing about accessibility in the lecture and then forgetting about it. 
Ideally, the assessments should not treat accessibility as a learning objective at a single point in the academic term. 
For instance, if accessibility is covered in Week 3 of the term, students can (and should) still be expected to write alternate texts for visualizations until the end of the semester.

@@

Although we have mainly focused on tools and different representations of data to have students develop accessible content, a more crucial accessibility feature is reproducible workflows. 
Scholars often focus on the reproducibility and replication of scientific findings. 
We believe reproducibility can often serve as an accessibility feature in a data science context. 
A reproducible open-source workflow, for example, allows users with and without dis/abilities to examine data in their preferred form of data representation.
Thus, in addition to teaching different representations of data, instructors may also consider teaching reproducibility.

@@

We believe that an accessible future in data science requires educators to teach accessibility as part of the curriculum deliberately. 
We also acknowledge that what we have presented here is only a small portion of the intersection of visual accessibility and data science.

Thus far, we have just demonstrated some methods to teach accessibility in introductory data science courses with a special focus on accessible visualization.
However, as @shinoharaWhoTeachesAccessibility2018 highlight, teaching accessibility at scale requires large-scale adoption and implementation of these methods by faculty, who are organically contributing to, and constructing, instructors’ knowledge [@barnettConceptualisingCurriculumChange2001; @kirkTeacherVoiceOwnership2001]. We hope that this manuscript has just opened the discussion on teaching accessibility in the data science field, and we encourage other data science faculty to help us move towards large-scale curricular change beyond our examples.

# Acknowledgements

The second co-author was supported through a Teach Access grant to develop curricular materials on teaching accessibility as part of the data science curriculum. 
Educators interested in teaching accessibility can follow the organization's work at https://teachaccess.org/.

# Supplementary Materials

The reproducible R source code of this manuscript can be found on a GitHub repository at \url{\repo{}}.

### 2022-11-13 revision_letter.Rmd
address reviewer comments

class       : "draft"
output      : papaja::revision_letter_pdf
bibliography: ["bib/packages.bib", "bib/references.bib"]
---

@@

# AE

@@

This is our response

\Assignment{First Author}
\WorkInProgress
\Easy

@@

\Assignment{Second Author}
\AlmostDone
\Medium

@@

This is our response: need to create a section dededicated to a11y, and say that we are presenting this paper as a demonstration how we integrated a11y into our courses.

\Assignment{First Author}
\Done
\TimeConsuming
\Hard
\NotEstimated

@@

This is our response

\Assignment{First Author}
\WorkInProgress
\Easy

@@

This is our response: good point so it is critical to share the same principle across different and with upper level courses .

\Assignment{First Author}
\WorkInProgress
\Easy

@@

This is our response

\Assignment{First Author}
\WorkInProgress
\Easy

>>>>

class       : "draft"
output:
  papaja::revision_letter_pdf:
    keep_tex: true
bibliography: ["bib/packages.bib", "bib/references.bib", "bib/teachinga11y.bib"]
csl: "bib/chicago-author-date.csl"
---

@@

# AE {#response1}

@@

We appreciate your thoughtful comment. In response to this feedback, we have elaborated our argument more clearly in our manuscript.

The following includes part of the `Introduction` section quoted from the revised manuscript to illustrate the change:

> Some people may wonder why we need to teach accessibility as part of an introductory data science course as opposed to covering accessibility in a separate and general course module. According to the recent faculty survey on teaching accessibility [@shinoharaWhoTeachesAccessibility2018], the most critical barriers to teaching accessibility reported by 1,857 faculty from 318 institutions were "the absence of clear and *discipline-specific accessibility* [emphasis added] learning objectives and the lack of faculty knowledge about accessibility" (p. 197). This reveals faculty often have trouble making sense of how they can integrate accessibility into their specific teaching domains, which requires applied knowledge of accessibility beyond general concepts.
> . . .
> Another important aspect of curriculum design includes the timing of when to introduce accessibility concepts in data science courses. For example, some instructors may prefer to put teaching accessibility on the back burner so that they can save more time for their students to learn data science fundamentals first and adopt accessibility principles in an upper-level course later. While the decision is subject to each institution's and individual instructor's discussion, we believe the sooner, the better. This is because: (1) some students may take an introductory level only, which means they will lose a chance to get exposed to this demanding knowledge unless they elect the upper-level course covering accessibility; (2) this kind of separation may unintentionally contribute to misconceptions, such as accessibility is difficult and/or accessibility is optional, not essential. When properly incorporated, accessibility can be taught along with other data science fundamentals (e.g., wrangling, merging, tidying, visualizing) while not necessarily putting more cognitive burden on students' current learning. Rather, students can take a chance to become creative design thinkers by making their data science work more meaningful and accessible through "empathy learning" [@putnamTeachingAccessibilityLearning2015]. Thus, we suggest teaching accessibility as part of an introductory data science course for students to practice coupling their data science knowledge with accessibility principles as early as possible.

@@

@@

Thank you for the constructive feedback. We strengthen our argument by adding the following paragraphs in the `Introduction` section.

> Data science consists of an iterative process of importing, tidying, transforming, visualizing, and modeling data [@wickham2016r]. While each stage has unique challenges, data visualization may pose fundamental barriers against people with permanent, situational, and temporary disabilities (e.g., visual, cognitive, and motor impairments, to name a few) in accessing information [@marriottInclusiveDataVisualization2021; @leeReachingBroaderAudiences2020]. For example, people with cognitive and learning disabilities often have difficulties understanding the abstractions and symbolic conventions used in visualizations, and people lacking fine motor coordination or having tremors may be unable to operate interactive visualizations, not to mention people who are blind and have low vision facing extra challenges in visualization access [@marriottInclusiveDataVisualization2021; @leeReachingBroaderAudiences2020]. @marriottInclusiveDataVisualization2021, in their call to action paper, highlights the serious impact of inaccessible visualizations: "This is an equity issue with severe consequences because a lack of access to visualizations and underlying data impacts health, educational and work opportunities, and lifestyle" (p. 48).
> Despite the increasing demand for accessibility knowledge in both industry and academia and the critical impact of accessible visualizations, accessibility has not yet found its place in the data science course and teaching materials. One main reason for this could be that accessibility is often incorporated into human-computer interaction (HCI) and design courses [@shinoharaHowDesigningPeople2016; @shinoharaTeachingAccessibilityTechnology2017a], introductory web development [@koAccessComputingPromotesTeaching2016; @rosmaitaAccessibilityFirstNew2006], software engineering courses [@martin-escalonaIntroductionTopicAccessibility2013; @ludiIntroducingAccessibilityRequirements2007; @ludiAccessEveryoneIntroducing2002] or the digital arts [@barata2019inclusion], if at all covered. However, data science is an interdisciplinary field that is taught in various disciplines, including not only computer and information sciences but also statistics, business schools, political science, biological sciences, and more [@schwab2021data]. Therefore, we posit that the resources and conversations around accessibility have not yet extended to the broader data science education community at the scale that they should. For instance, many introductory data science courses [@yan2019first; @baumer2015data], textbooks [@adhikari2019computational; @wickham2016r], or curricula [@schwab2021data] either fail to mention accessibility at all or only focus on accessing the data.
> In response to the need to develop accessible teaching materials within the data science discipline, this paper demonstrates and shares some teaching strategies that we have employed in our teaching introductory data science courses.

@@

Thanks for the insightful comment. We have addressed this question as described in our responses to AE's comments above.

@@

Thank you very much for your great input. Although we demonstrated our examples in the context of introductory data science courses, we believe accessibility needs to be taught in other data science courses, including higher levels. This requires instructors' broader efforts. We have clarified this in our `Closing Remarks` section like below:

> Thus far, we have just demonstrated some methods to teach accessibility in introductory data science courses with a special focus on accessible visualization.
> However, as @shinoharaWhoTeachesAccessibility2018 highlight, teaching accessibility at scale requires large-scale adoption and implementation of these methods by faculty, who are organically contributing to, and constructing, instructors’ knowledge [@barnettConceptualisingCurriculumChange2001; @kirkTeacherVoiceOwnership2001]. We hope that this manuscript has just opened the discussion on teaching accessibility in the data science field, and we encourage other data science faculty to help us move towards large-scale curricular change beyond our examples.

@@

We have addressed this comment above via [our first response to AE's comment](#response1).
